---
---
---

# Capstone Project

Author: Ismail Khalil\
Date: "2022-11-01"\
Output: github_document

## Summary

You are a junior data analyst working for a business intelligence consultant. You have been at your job for six months, and your boss feels you are ready for more responsibility. He has asked you to lead a project for a brand new client---this will involve everything from defining the business task all the way through presenting your data-driven recommendations. You will choose the topic, ask the right questions, identify a fresh dataset and ensure its integrity, conduct analysis, create compelling data visualizations, and prepare a presentation.

## Ask

Business Task: My client is the executive team from Nintendo, and they want me to analyze video game trends in the past to produce insights that will help them with their future products.

The dataset I will use for my analysis will be from Kaggle, and it is a dataset about video game sales greater than 100,000 copies from 1980 to 2016.

A link to the dataset can be found by [clicking here](https://www.kaggle.com/datasets/gregorut/videogamesales?pag=).\

## Prepare

One of the limitations of this dataset is that it only goes up to 2016, and is not current. For the purposes of this project only this dataset will be used, but for a full analysis more current data is needed.

The data is organized in long format and there are no bias or credibility issues as the full process on how they got the data was detailed in their Kaggle description.

This data will help me derive insights because it has sales data related to a large range of companies as well as Nintendo games.

The data from vgchartz has good data integrity, but the data are estimates.

## Process

To start I will download the relevant libraries, load in the dataset, and then clean up the data that I choose to use for my analysis.

```{r}
install.packages('tidyverse')
library('tidyverse')
```

```{r}
sales_data_df <- read_csv("vgsales.csv")
```

```{r}
view(sales_data_df)
```

The sales columns are the number of copies sold in millions and not the revenue.

```{r}
glimpse(sales_data_df)
```

Cleaning the data to make sure it only has games from 1980 to 2016

```{r}
sales_data_df2 <- sales_data_df %>%
filter(sales_data_df$Year >= 1980, sales_data_df$Year <= 2016)
```

Each variable does not have any missing values

```{r}
sapply(sales_data_df2, function(x) sum(is.na(x)))
```

After using the unique function as well as carefully looking through each of the columns, the dataset is now clean.

## Analyze & Share

I want to first find out which publisher had the highest global sales out of all the years in this data set

```{r}
global_sales_by_pub <- aggregate(sales_data_df2$Global_Sales, list(sales_data_df2$Publisher), FUN=sum)
colnames(global_sales_by_pub) <- c("Publisher", "Global_Sales")
```

The top five publishers in global sales from 1980 to 2016 are:

1.  Nintendo
2.  Electronic Arts (EA)
3.  Activision
4.  Sony Computer Entertainment
5.  Ubisoft

I now want to find out which games in particular earned the highest global sales

```{r}
global_sales_by_game <- (aggregate(sales_data_df2$Global_Sales, list(sales_data_df2$Name), FUN=sum))
colnames(global_sales_by_game) <- c("Name", "Global_Sales")
```

It looks like the top five games for this time period is:

1.  Wii Sports
2.  Grand Theft Auto V
3.  Super Mario Bros.
4.  Tetris
5.  Mario Cart Wii

All of these games were published by Nintendo except for Grand Theft Auto which was published by Take Two Interactive.

Now I want to see what genres are potentially on the rise by looking at global sales

```{r}
ggplot(sales_data_df2, aes(x = Genre, y = Global_Sales,fill = Genre, color = Genre)) + geom_bar(stat = "identity") + coord_flip() + ggtitle( 'Global Sales per Genre')

```

```{r}
ggplot(sales_data_df2, aes(x=Genre))+ geom_bar() + labs(y = "Games Released per Genre")+ coord_flip()+ ggtitle("Number of Games Released per Genre")
```

The genre with the highest number of sales was action. However, the highest number of games released during this dataset were also under the action genre, which means that it is inconclusive whether or not action games are on the rise because we do not know whether or not it was the action genre that caused people to buy more of it or that there were so many action games games available that people just chose to buy those instead if games from other genres were not available. In fact, the number of games released per genre look proportional to the global sales per genre. It could be the case that because action games released so many more games they had more chances of having a game be a hit, which might have been the reason why they have the highest number of sales.

I now want to take a look at all the all the games Nintendo released and the genres that each of them were in.

```{r}
nintendo_games_released <- sales_data_df2 %>% filter(Publisher== "Nintendo")
ggplot(nintendo_games_released, aes(x=Genre)) +geom_bar() + labs(y="Number of Games Released") + coord_flip() + ggtitle("Nintendo Games Released by Genre")
```

Nintendo has produced the most Platform games and the least Fighting games. Nintendo should increase the number of Shooter games that they are releasing because if you look at global sales then you can see that that is one of the most popular genres, but Nintendo is not making enough of those games. Nintendo has the possibility of taking more market share away from the games currently producing Shooter games.

I now want to see the top five publishers in the shooting genre and their global sales.

```{r}
shooter_games_released <- sales_data_df2 %>% filter(Genre== "Shooter" & (Publisher == "Nintendo" |Publisher == "Electronic Arts" |Publisher == "Activision" |Publisher == "Ubisoft" |Publisher == "Microsoft Game Studios" ))
ggplot(shooter_games_released, aes(x = Publisher, y = Global_Sales, fill = Publisher, color = Publisher)) + geom_bar(stat = "identity") + ggtitle("Global Sales per Publisher in Shooting Genre") + coord_flip()
```

It is clear that Nintendo is lagging far behind in terms of sales in the shooting genre. If they want to keep up with the rest of the publishers in this genre they need to create games that could possibly be on par.

## Act

To conclude, I have three insights I want to share with the Nintendo executive team: Nintendo is the number one publisher in terms of global sales from 1980 to 2016. The game that earned the highest global sales in this time period was Wii Sports, which means that Nintendo should possibly publish a game similar to that in the future. Lastly, Nintendo should stay out of the shooting genre unless they can create a game that is on par with the likes of games, such as Activision's Call of Duty franchise.
